---
title: "Reproducible Research: Peer Assessment 1"
author: "Kleanthis Mazarakis"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
activity_data <- read.csv("activity.csv")
head(activity_data)
str(activity_data)
activity_data$dateTime <- as.Date(activity_data$date, format = "%Y-%m-%d")
```

## What is mean total number of steps taken per day?
```{r}
StepsPerDay <- tapply(activity_data$steps, activity_data$dateTime, FUN = "sum")
hist(StepsPerDay)

mean_per_day <- mean(StepsPerDay, na.rm = T)
mean_per_day
median_per_day <- median(StepsPerDay, na.rm = T)
median_per_day
```
The mean steps per day was 10766 and the median was 10765

## What is the average daily activity pattern?
```{r}
interval_average <- tapply(activity_data$steps, activity_data$interval, mean, na.rm = T)
intervals_factor <- factor(activity_data$interval)
plot(levels(intervals_factor), interval_average, type = "l", xlab = "Intervals", ylab = "Average Steps", main = "Average Daily Activity Pattern")
```

Now let's check which 5 min interval, on average, contains the maximum number of steps
```{r}
# Creating a data frame with the intervals and the average steps per interval and then will select the maximum row
df <- data.frame(levels(intervals_factor), interval_average)
df[df$interval_average == max(df$interval_average),]
```
So the maximum number of steps per interval is 206 and it occured on interval 835.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
